---
title: "Bivariate Viz"
author: Feride Karatas
---

this file to generate a professional looking **bivariate** visualization. The visualization will not perfect the first time but you are expected to improve it throughout the semester especially after covering advanced topics such as effective viz.

---
title: "Bivariate Viz"
number-sections: true
execute: 
  warning: false
fig-height: 2.75
fig-width: 4.25
fig-env: 'figure'
fig-pos: 'h'
fig-align: center
code-fold: false
---

# 

```{r}
# Load necessary libraries
library(ggplot2)
library(tidyverse)


elections <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")


head(elections)

```

```{r}

ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point(color = "blue", alpha = 0.6) +  # Scatter plot points
  geom_smooth(method = "lm", color = "red", se = TRUE) +  # Linear trendline with confidence interval
  labs(
    title = "Bivariate Graph ofrepub_pct_16 vs. repub_pct_20",
    x = "repub_pct_16",
    y = "repub_pct_20",
    caption = "Elections"
  ) +
  theme_minimal() +
  theme(text = element_text(size = 12))  # Improve readability

```
