---
title: "Trivariate Viz"
---

Use this file to generate a professional looking **trivariate** visualization.
The visualization will not perfect the first time but you are expected to improve on it throughout the semester especially after covering advanced topics such as effective viz.

```{r}
library(ggplot2)
library(tidyverse)
```

```{r}

# Code to create the scatter plot
library(ggplot2)
library(palmerpenguins)

ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g, color = species, shape = sex)) +
  geom_point() +
  labs(
    title = "Penguin Body Mass vs Bill Length by Species and Sex",
    x = "Bill Length (mm)",
    y = "Body Mass (g)"
  ) +
  theme_minimal()


```

> The scatter plot is effective in visualizing the relationship between penguin bill length and body mass, differentiated by species and sex.
> The use of color (species) and shape (sex) helps separate the groups clearly without overwhelming the plot.
>
> We observe that Gentoo penguins tend to have longer bills and heavier body masses compared to Adelie and Chinstrap penguins.
> Additionally, males (different shapes) generally have larger measurements compared to females within each species.
>
> This visualization is particularly effective for identifying group trends and overlaps between species and sexes at the same time.
> If we were more interested in comparing only averages, a grouped boxplot would be more appropriate but for exploring individual variation and patterns, this scatter plot works better.
