---
title: "Exam 1"
format: html
---

**Food Consumption & CO₂ Emissions**

**Research Question:** What does the consumption of each food category in each country look like?

```{r}
#| code-fold: true
#| warning: false
#| error: false
#| echo: true
#| results: "show"
library(tidytuesdayR)
library(tidyverse)

```

```{r}
#| code-fold: true
#| warning: false
#| error: false
#| echo: true
#| results: "show"
tuesdata <- tt_load('2020-02-18')
fc <- tuesdata$food_consumption

```

```{r}
#| code-fold: true
#| warning: false
#| error: false
#| echo: true
#| results: "show"
head(fc)      # view first few rows
tail(fc)      # view last few rows
nrow(fc)      # number of rows
ncol(fc)      # number of columns
class(fc)     # class of the object

```

```{r}
#| code-fold: true
#| warning: false
#| error: false
#| echo: true
#| results: "show"
head(fc, 22)

```

```{r}
#| code-fold: true
#| warning: false
#| error: false
#| echo: true
#| results: "show"
tail(fc, 22)

```

```{r}
#| code-fold: true
#| warning: false
#| error: false
#| echo: true
#| results: "show"
#| fig-height: 8
#| fig-width: 12

ggplot(fc, aes(x = food_category, y = consumption)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Food Consumption per Food Category",
    x = "Food Category",
    y = "Consumption (kg/person/year)"
  )

```

```{r}
#| code-fold: true
#| warning: false
#| error: false
#| echo: true
#| results: "show"
#| fig-height: 22
#| fig-width: 11

ggplot(fc, aes(x = food_category, y = consumption, fill = food_category)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ country, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Food Category Consumption by Country",
    x = "Food Category",
    y = "Consumption (kg/person/year)"
  )

```

```{r}
#| code-fold: true
#| warning: false
#| error: false
#| echo: true
#| results: "show"
#| fig-height: 22
#| fig-width: 11

ggplot(fc, aes(x = food_category, y = consumption, fill = food_category)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ country) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Food Consumption by Category and Country",
    x = "Food Category",
    y = "Consumption (kg/person/year)"
  )

```

> Throughout this project, I strengthened my skills in loading data, performing exploratory analysis, and choosing appropriate visualizations.
> Managing the size of plots was important to make sure the faceted bar charts were readable.
> One challenge was handling missing values and making the x-axis labels readable by rotating them.
> If I did this again, I would also add a choropleth map to explore the geographical patterns of food consumption more visually.
> This project helped me see the importance of clear labeling, accessibility, and not misleading the viewer.

