---
format: html
---

# Solo Project

## Exploring Average Income Across the State of California

### Information:

> The dataset contains data on test performance, school characteristics and student demographic backgrounds for school districts in California.
>
> Check the [codebook](https://vincentarelbundock.github.io/Rdatasets/doc/AER/CASchools.html) here.
>
> #### Details
>
> The data used here are from all 420 K-6 and K-8 districts in California with data available for 1998 and 1999.
> Test scores are on the Stanford 9 standardized test administered to 5th grade students.
> School characteristics (averaged across the district) include enrollment, number of teachers (measured as “full-time equivalents”, number of computers per classroom, and expenditures per student. Demographic variables for the students are averaged across the district. The demographic variables include the percentage of students in the public assistance program CalWorks (formerly AFDC), the percentage of students that qualify for a reduced price lunch, and the percentage of students that are English learners (that is, students for whom English is a second language).
>
> In this solo project I wanted to see if there is variation in income across the state of California which is known by how expensive it is especially by the coast.
> This project can be expanded by analyzing whether income has effect on math, reading etc grades and the addition of confounders such as the amount of students per class to ensure no misinformation is being made and the true relationship is explored.
>
> ### References
>
> Stock, J. H.
> and Watson, M. W.
> (2007).
> *Introduction to Econometrics*, 2nd ed.
> Boston: Addison Wesley.

```{r}
library(sf)
library(tidyverse)
library(ggplot2)
library(mosaic)
testscore <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/AER/CASchools.csv")
cal_boundries <- read_sf("../data/California_County_Boundaries_8408091426384550881/cnty19_1.shp")

ggplot(cal_boundries)+
  geom_sf()
```

```{r}

# Prepare test score data by averaging income per county
testscore_summary <- testscore %>%
  group_by(county) |>
  summarize(income = mean(income, na.rm = TRUE))

# Join with shapefile data
cal_boundries <- cal_boundries %>%
  left_join(testscore_summary, by = c("COUNTY_NAM" = "county"))

# Plot Choropleth
ggplot(cal_boundries) +
  geom_sf(aes(fill = income)) +
  scale_fill_viridis_c(option = "plasma") +  # Better visualization
  theme_minimal() +
  labs(title = "Choropleth of Income by County",
       fill = "Average Income")+
  theme_map()


```

```{r}

# Prepare test score data by averaging income per county
testscore_summary <- testscore %>%
  group_by(county) |>
  summarize(math = mean(math, na.rm = TRUE))

# Join with shapefile data
cal_boundries <- cal_boundries %>%
  left_join(testscore_summary, by = c("COUNTY_NAM" = "county"))

# Plot Choropleth
ggplot(cal_boundries) +
  geom_sf(aes(fill = math )) +
  scale_fill_viridis_c(option = "plasma") +  # Better visualization
  theme_minimal() +
  labs(title = "Choropleth of Math Scores by County",
       fill = "Average Math Scores")+
  theme_map()

```

## Conclusion:

> San Jose and San Francisco area has people with the highest average income.
> This could be due to San Francisco and San Jose being very expensive and only people who earn a certain threshold of income can live in those regions.
>
> I have also included a choropleth plot of average math scores across California and visually it is similar to that of average income.
> Meaning that there might be a potential relationship between the two variables.
> The relationship can be explored with a scatterplot accounting for effect modifiers and confounders.
> I find this data interesting and it makes sense for me.
