---
title: "Reshaping"
number-sections: true
execute: 
  warning: false
fig-env: 'figure'
fig-pos: 'h'
fig-align: center
code-fold: false
---

::: {.callout-caution title="Learning Goals"}
-   Understand what it means to "reshape" data.
-   Understand the difference between *wide* and *long* data formats.
-   Be able to distinguish the units of observation for a given data set.
-   Explore how to reshape data using `pivot_wider()` and `pivot_longer()` from `tidyr` package
:::

::: {.callout-note title="Additional Resources"}
For more information about the topics covered in this chapter, refer to the resources below:

-   [Demonstrating pivoting (YouTube)](https://www.youtube.com/watch?v=k3SZ8keibuQ&feature=youtu.be) by Lisa Lendway
-   [Pivoting vignette (html)](https://tidyr.tidyverse.org/articles/pivot.html) by tidyr
-   [Pivoting (html)](https://r4ds.hadley.nz/data-tidy.html#sec-pivoting) by Wickham and Grolemund
-   [Reshaping data](https://mdsr-book.github.io/mdsr2e/ch-dataII.html#reshaping-data) by Baumer, Kaplan, and Horton

## 
:::

```{r}
library(tidyverse) 
penguins <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-28/penguins.csv')

penguins |> count(species, sex)

penguins |> group_by(species, sex) |> summarize(n())

penguins |> group_by(species, sex) |> summarize(n = n()) |> mutate(proportion = n / sum(n))
```

```{r}
penguin_avg <- penguins |> 
  group_by(species, sex) |> 
  summarize(avg_body_mass = mean(body_mass_g, na.rm = TRUE)) |> 
  na.omit()

# Units of observation = penguins
head(penguins)

# Units of observation = species/sex combos
head(penguin_avg)
penguin_avg


```

```{r}
penguin_avg |> 
  pivot_wider(names_from = sex, values_from = avg_body_mass)

```

```{r}
penguin_avg_wide <- penguin_avg |> 
  pivot_wider(names_from = sex, values_from = avg_body_mass)

penguin_avg_wide

# Pivot longer again
penguin_avg_wide |> 
  pivot_longer(cols = c(female, male), names_to = "sex", values_to = "avg_body_mass")

```

```{r}
food <- data.frame(
  customer = rep(c("A", "B"), each = 3),
  restaurant = rep(c("Shish", "FrenchMeadow", "DunnBros"), 2),
  order = c("falafel", "salad", "coffee", "baklava", "pastry", "tea")
)

# Make wide
food |> pivot_wider(names_from = restaurant, values_from = order)

more_food <- data.frame(
  customer = c("C", "D"),
  Shish = c("coffee", "maza"),
  FrenchMeadow = c("soup", "sandwich"),
  DunnBros = c("cookie", "coffee")
)

# Make long
more_food |> pivot_longer(cols = -customer, names_to = "restaurant", values_to = "order")

```
