---
title: "ica-dataimp"
format: html
editor: 
  markdown: 
    wrap: 72
---

---
title: "Data Import"
number-sections: true
execute: 
  warning: false
fig-env: 'figure'
fig-pos: 'h'
fig-align: center
code-fold: false
---

::: {.callout-caution title="Learning Goals"}
Get a glimpse into how to... - find existing data sets - save data sets
locally - load data into RStudio - do preliminary data checking and
cleaning: - format variables properly - deal with missing values NOTE:
Best learned through practice.
:::

::: {.callout-note title="Additional Resources"}
-   [Import wizard demo
    (YouTube)](https://www.youtube.com/watch?v=GtCsjtZBNp4)
-   [Data Import Cheat Sheet
    (pdf)](https://github.com/rstudio/cheatsheets/raw/main/data-import.pdf)
-   [`readr` documentation](https://readr.tidyverse.org/)
-   [Data Import (r4ds)](https://r4ds.hadley.nz/data-import.html)
-   [Missing data (r4ds)](https://r4ds.hadley.nz/missing-values.html)
-   [Data intake
    (MDSR)](https://mdsr-book.github.io/mdsr2e/ch-dataII.html#data-intake)
:::

```{r}


library(tidyverse)

# Import the messy IMDB dataset from Downloads
imdb_messy <- read_csv("~/Desktop/imdb_5000_messy.csv")


str(imdb_messy)

head(imdb_messy)

# Summarize data
imdb_messy |>
  mutate(across(where(is.character), as.factor)) |>
  summary()

```

```{r}
# See how color variable looks
imdb_messy |> count(color)

# Clean it using fct_recode()
imdb_clean <- imdb_messy |> 
  mutate(color = fct_recode(color,
                            "Color" = "COLOR",
                            "Color" = "color",
                            "Black_White" = "B&W",
                            "Black_White" = "Black and White"))

# Double-check cleaned values
imdb_clean |> count(color)

```

```{r}
# How many missing values per column?
colSums(is.na(imdb_messy))

# Look at movies missing actor 1 facebook likes
imdb_messy |> 
  filter(is.na(actor_1_facebook_likes))

```

```{r}
# Drop movies missing duration or color
imdb_no_missing <- imdb_messy |> 
  filter(!is.na(duration), !is.na(color))

# Plot to check
ggplot(imdb_no_missing, aes(x = duration, fill = color)) +
  geom_density()

```

